---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# testinterference: Randomization Test for the Specification of Interference Structure

<!-- badges: start -->
<!-- badges: end -->

The **testinterference** package provides tools to test SUTVA and several hypotheses about spillover effects.
Specifically, the package enables to perform randomization tests on whether or not the following null hypotheses are plausible:

- **Fisher**:
One's outcome does not depend on one's own treatment (i.e., Fisher's sharp null hypothesis).
- **SUTVA**:
One's outcome is determined by one's own treatment (i.e., the stable unit treatment value assumption).
- **Exposure 1**:
One's outcome is determined by whether there is at least one treated unit in one's neighborhood including oneself.
- **Exposure 2**:
One's outcome is determined by one's own treatment and whether there is at least one treated peer.

The testing procedures are developed by [Hoshino and Yanagi (2023) "Randomization test for the specification of interference structure"](https://doi.org/10.48550/arXiv.2301.05580).

## Installation

Get the package from GitHub:

```{r, eval = F, echo = T}
# install.packages("devtools") # if needed
devtools::install_github("tkhdyanagi/testinterference", build_vignettes = TRUE)
```

## Vignettes

For more details, see the package vignettes with:

```{r, eval = FALSE}
# Getting Started with the testinterference Package
vignette("testinterference")

# Testing Procedure: Fisher's Sharp Null Hypothesis
vignette("Fisher", package = "testinterference")

# Testing Procedure: SUTVA
vignette("SUTVA", package = "testinterference")

# Testing Procedure: Exposure 1
vignette("exposure1", package = "testinterference")

# Testing Procedure: Exposure 2
vignette("exposure2", package = "testinterference")
```

## References

- Hoshino, T. and Yanagi, T., 2023. 
Randomization test for the specification of interference structure
arXiv preprint arXiv:2301.05580.
[Link](https://doi.org/10.48550/arXiv.2301.05580)
