---
title: "Getting Started with the testinterference Package"
author: "Tadao Hoshino (thoshino@waseda.jp), Takahide Yanagi (yanagi@econ.kyoto-u.ac.jp)"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started with the testinterference Package}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7.5, 
  fig.height = 5 
)
options(rmarkdown.html_vignette.check_title = FALSE)
```

## Introduction

The **testinterference** package provides tools to test SUTVA and several hypotheses about spillover effects.
Specifically, the package enables to perform randomization tests on whether or not the following null hypotheses are plausible:

- **Fisher**:
One's outcome does not depend on one's own treatment (i.e., Fisher's sharp null hypothesis).
- **SUTVA**:
One's outcome is determined by one's own treatment (i.e., the stable unit treatment value assumption).
- **Exposure 1**:
One's outcome is determined by whether there is at least one treated unit in one's neighborhood including oneself.
- **Exposure 2**:
One's outcome is determined by one's own treatment and whether there is at least one treated peer.

The testing procedures are developed by [Hoshino and Yanagi (2023) "Randomization test for the specification of interference structure"](https://doi.org/10.48550/arXiv.2301.05580).

## Installation

Get the package from GitHub:

```{r, eval = F, echo = T}
# install.packages("devtools") # if necessary
devtools::install_github("tkhdyanagi/testinterference", build_vignettes = TRUE)
```

## Package Function

The **testinterference** package provides the following function:

- `testinterference()`: Randomization tests for the following null hypotheses: (1) Fisher, (2) SUTVA, (3) Exposure 1, (4) Exposure 2.

#### Arguments

The `testinterference()` function has the following arguments:

- `Y`: The $n$-dimensional outcome vector
- `Z`: The $n$-dimensional binary treatment assignment vector
- `A`: The $n \times n$, binary, possibly asymmetric, adjacency matrix.
   The diagonal elements must be zero.
   This argument can be NULL if `hypothesis = Fisher`.
   Default is NULL.
- `hypothesis`: A character specifying the null hypothesis of interest.
   Options are "Fisher", "SUTVA", "exposure1", and "exposure2".
   "Fisher" stands for the sharp null hypothesis in the canonical Fisher randomization test.
   "SUTVA", "exposure1", and "exposure2" correspond to 
   the null hypotheses a, b, and c in Table 2 of Hoshino and Yanagi (2023).
   Default is "SUTVA".
   If `hypothesis = "exposure2"`, the user can specify the argument `kappa`.
- `method`: A character specifying how to find focal units.
   Options are "MIS", "random", and "manual", 
   which stand for the method of maximum independent set (MIS), the random selection, and the manual selection.
   Default is "MIS".
   If `method = "random"`, the user can specify the argument `num_focal_unit`.
   If `method = "manual"`, the user must specify the argument `focal_unit`.
- `design`: A character specifying the randomization design of the original experiment.
   Options are "Bernoulli", "complete", "stratified", and "other", 
   which stand for Bernoulli randomization, complete randomization, stratified randomization, and other experimental designs.
   If `design = "Bernoulli"`, the user must specify the argument `prob`.
   If `design = "stratified"`,the user must specify the argument `strata`.
   If `design = "other"`, the user must specify the argument `zmatrix`.
- `prob`: NULL, a scalar indicating the homogeneous probability of being treatment eligible,
   or an $n$-dimensional vector indicating the heterogeneous probabilities
   for treatment eligibility (i.e., the $i$-th element of `prob` indicates $\Pr(Z_i = 1)$).
   This argument is used only when `design = "Bernoulli"`.
   Default is NULL.
- `focal_unit`: NULL or an $n$-dimensional logical vector specifying whether each unit is focal.
   This argument is used only when `method = "manual"`.
   Default is NULL.
- `num_focal_unit`: NULL or a positive scalar specifying the number of focal units.
   This argument is used only when `method = "random"`.
   Default is NULL.
- `num_randomization`: A large positive integer specifying the number of randomization.
   Default is 9999.
   This argument is not used when `method = "other"`.
- `strata`: NULL or an $n$-dimensional numerical vector indicating the stratum to which each unit belongs.
   This argument is used only when `design = "stratified"`.
   Default is NULL.
- `zmatrix`: NULL or a large matrix of realizations of treatment assignments.
   This argument must be specified by the user only when `design = "other"`.
   The number of rows must equal to the sample size $n$.
   The number of columns is the number of realizations given by the user.
   Default is NULL.
- `kappa`: NULL or a positive integer no less than 2.
   This argument is used only when `hypothesis = "exposure2"`.
   Default is NULL.
   If `kappa = NULL`, kappa is automatically chosen to maximize the number of focal units.
- `cores`: A positive integer specifying the number of cores to use for parallel computing.
   Default is 1.

#### Returns

The `testinterference()` function returns a list containing the following elements:

- `pval`: The vector of p-values obtained from Kruskal-Wallis (KW), average cross difference (ACD), and ordinary least squares (OLS) test statistics.
- `Simes`: The vector of testing results by Simes' correction under significance levels 10%, 5%, and 1%.
   TRUE (resp. FALSE) indicates the rejection (resp. acceptance) of the null hypothesis.
- `stat`: A matrix of test statistics computed with focal assignments.
- `focal_unit`: The logical vector indicating whether each unit is focal.
- `focal_asgmt`: The matrix of focal assignments.
- `num_focal_unit`: The number of focal units.
- `num_focal_asgmt`: The number of focal assignments.

## Example 1: A Single Large Network 

We begin by generating artificial data using the `datageneration()` function.
In the `datageneration()` function, there are two options to generate the adjacency matrix: 
(1) Erdos-Renyi model and (2) pairs.
In addition, there are three options for the experimental design: 
(a) Bernoulli randomization, (b) complete randomization, and (c) stratified randomization.
Here we consider the case (1)-(a) with:

```{r, eval = T, echo = T}
# Load the package
library(testinterference)

# Sample size
n <- 200

# Generate artificial data (Erdos-Renyi model)
set.seed(1)
data1 <- datageneration(n      = n,
                        design = "Bernoulli",
                        A      = "Erdos-Renyi",
                        beta_s = 1)
```

Here, `beta_s` specifies the value of a coefficient for spillover effects in the outcome equation.
`beta_s = 0` means that there are no spillovers.
In this example, we consider the case where SUTVA does not hold.

Run the `testinterference()` function to test SUTVA:

```{r, eval = T, echo = T}
set.seed(1)
RT1 <- testinterference(Y                 = data1$Y,
                        Z                 = data1$Z,
                        A                 = data1$A,
                        hypothesis        = "SUTVA",
                        method            = "MIS",
                        design            = "Bernoulli",
                        prob              = 0.5,
                        focal_unit        = NULL,
                        num_focal_unit    = NULL,
                        num_randomization = 999,
                        strata            = NULL,
                        zmatrix           = NULL,
                        kappa             = NULL,
                        cores             = 1)
```

Here, we set `num_randomization = 999` to reduce the computation time, 
but in realistic situations the number of randomization should be larger (e.g., `num_randomization = 99999`).

The p-values obtained from Kruskal-Wallis (KW), average cross difference (ACD), and ordinary least squares (OLS) test statistics are:
```{r, eval = T, echo = T}
RT1$pval
```

The null hypothesis is rejected at the standard significance level.
In other words, we find statistical evidence that SUTVA is implausible.

The results of Simes' correction are:

```{r, eval = T, echo = T}
RT1$Simes
```

Here, `TRUE` indicates the rejection of the null hypothesis.
Again, the results suggest that SUTVA does not hold.

We can confirm the numbers of focal units and assignments with:

```{r, eval = T, echo = T}
RT1$num_focal_unit
RT1$num_focal_asgmt
```

Next, we turn to test whether exposure 2 is correct.

```{r, eval = T, echo = T}
set.seed(1)
RT2 <- testinterference(Y                 = data1$Y,
                        Z                 = data1$Z,
                        A                 = data1$A,
                        hypothesis        = "exposure2",
                        method            = "MIS",
                        design            = "Bernoulli",
                        prob              = 0.5,
                        focal_unit        = NULL,
                        num_focal_unit    = NULL,
                        num_randomization = 999,
                        strata            = NULL,
                        zmatrix           = NULL,
                        kappa             = NULL,
                        cores             = 1)
```

The testing results are:
```{r, eval = T, echo = T}
RT2$pval
RT2$Simes
RT2$num_focal_unit
RT2$num_focal_asgmt
```

We find statistical evidence that exposure 2 is incorrect.

## Example 2: Pairs

We turn to the case with possible pairwise interactions and complete randomization.
To consider the case without interference, we specify `beta_s = 0`.

```{r, eval = T, echo = T}
# Generate artificial data (Erdos-Renyi model)
set.seed(1)
data2 <- datageneration(n      = n,
                        design = "complete",
                        A      = "pairs",
                        beta_s = 0)
```

Perform the randomization test:

```{r, eval = T, echo = T}
set.seed(1)
RT3 <- testinterference(Y                 = data2$Y,
                        Z                 = data2$Z,
                        A                 = data2$A,
                        hypothesis        = "SUTVA",
                        method            = "MIS",
                        design            = "complete",
                        prob              = NULL,
                        focal_unit        = NULL,
                        num_focal_unit    = NULL,
                        num_randomization = 999,
                        strata            = NULL,
                        zmatrix           = NULL,
                        kappa             = NULL,
                        cores             = 1)
```

The testing results are:

```{r, eval = T, echo = T}
RT3$pval
RT3$Simes
RT3$num_focal_unit
RT3$num_focal_asgmt
```

The null hypothesis is not rejected at the standard significance level.
For the results of Simes' correction, `FALSE` indicates the acceptance of the null hypothesis.
Thus, the testing results suggest that SUTVA might hold.

In the case with pairwise interactions, 
it is impossible to perform randomization tests for `hypothesis = "exposure2"`.
This is because each unit has only one peer so that $\mathcal{N}(\kappa)$ is empty for any $\kappa \ge 2$.
See the vignette `vignette("exposure2", package = "testinterference")` for more details.

## References

- Hoshino, T. and Yanagi, T., 2023. 
Randomization test for the specification of interference structure
arXiv preprint arXiv:2301.05580.
[Link](https://doi.org/10.48550/arXiv.2301.05580)
